---
title: "Earthquake!"
author: "Lauren Ng"
date: "12/2/2019"
output: html_document
---

```{r echo=FALSE, messages=FALSE}
source("earthquake_analysis.R")
```

## Introduction
Earthquakes are a force to be reckoned with. They’re powerful jolts caused by
movement underneath the surface that can drastically change our world and
environment. For this project, I will be analyzing a database that records
significant earthquakes  (5.5+ magnitude) from 1965 to 2016. The entire
dataset is compiled by The National Earthquake Information Center (NEIC).
Which takes data from national and international agencies, scientists,
critical facilities, and the general public and complies the information
gathered. The main goal of this dataset is to provide another resource for
scientists researching trends and networks of global seismographic reports.
But also utilized for public knowledge and interpretations since the source
is on kaggle and accessible to everyone. [Here is the link to the database](https://www.kaggle.com/usgs/earthquake-database)

## Data + Discussion 
Looking at the various plots of the graph can display a lot of information
that can often be looked over or ignored when looking just at the dataset,
since it is very dense. Using plots can summarize the data collected and make
the data visualizable. Since the information was very dense, I have made a
couple plots to understand most aspects of the dataset.

```{r echo=FALSE, messages=FALSE}
year_breakdown 
```

The first visualization is to answer the question: Are the number of
earthquakes increasing every year? To answer, I added the number of
earthquakes every year and plotted the values on a line graph. This enables
readers to see the trend of number of earthquakes each year.  From first
glance, it looks like the number of earthquakes fluctuates randomly where
one year is 100 more or less than the previous year. However, looking
carefully, there is an upward trend. Which might mean that future years might
have a large amount of earthquakes each year. And this could affect more than
the people living on the fault line. However, we must also ask the question
of whether this information is reliable. The technology in the 1960’s might
not have detected every significant earthquake due to the lack of sophisticated
technology, so it is up to us to decide if it’s accurate. Nevertheless, there
is always around 500 earthquakes each year so citizens be warned!

```{r echo=FALSE, messages=FALSE}
mapping
```
caption: displays location of earthquakes 

This plot displays the locations of earthquakes, which is important when
trying to understand where to be more cautious and predict future locations
of earthquakes. Back then, researchers were able to determine the location
of fault lines by these earthquake recordings. So highlighting these locations
might uncover a new fault line or discovery. It also informs citizens in the
areas to be prepared for earthquakes since there has been a history of
earthquakes. 

## Reflection 
Overall, this extra analysis has pushed me to learn more about R and about a
topic that has impacted my life. Growing up in California, we regularly had
earthquakes, both small and large. Understanding the reasons why and when can
be really beneficial since as a child, I thought the world was going to crumble
when an earthquake hit. It was a really neat experience to relate data science
to something I care about and learn more about the topic. I also feel like I
learned a lot more about R and practiced the skills I learned in the course on
my own. There is much more that meets the surface, both technically and
information based. There is always more to the information presented, like we
saw here with the dataset and visualizations. And technically, especially when
coding in R. We only hit the surface but there is a lot more that we haven’t
explored. And doing this project has encouraged me to take on my own projects,
since I was able to do this project without much help. 

## My code for the data analysis 
df <- read.csv("database.csv")
### Year breakdown
year <- df %>%
  mutate(year = as.numeric(substring(df$Date, 7,11))) %>%
  group_by(year) %>%
  count()
officalYear <- year[complete.cases(year), ]

year_breakdown <- ggplot(officalYear, aes(x=year, y=n)) +
  scale_x_continuous(breaks=seq(1965, 2016, 5)) + 
  geom_line(group = 1) + labs(x = "year", y = "# of earthquakes",
                              title = "Number of Earthquakes every year",
                              caption = "(based on data from earthquake database. 
                              which only counts significant earthquakes that have 
                              a magnitude of 5.5 or higher.)")

### Map of earthquake locations
map <- list(
  scope = "world",
  showland = TRUE,
  landcolor = toRGB("gray95"),
  countrycolor = toRGB("black")
)

map <- plot_geo(df, lat = ~Latitude, lon = ~Longitude) %>%
add_markers(
  text = ~paste("State: ", Magnitude),
  colors = c("yellow", "orange", "red", "red2", "red3"),
  color = ~Magnitude, symbol = I("circle"), hoverinfo = "text",
  marker = list(size = ~Magnitude)) %>%
  colorbar(title = "Magnitude") %>%
  layout(title = "Locations of Earthq", geo = map)